% -*- mode: LaTeX; TeX-PDF-mode: t; -*-  # Config emacs auctex

% allow latex to find custom stuff
\input{@resources/tex-add-search-paths}  

\documentclass[12pt,letterpaper,english]{article}
\usepackage[round, authoryear]{natbib}
\usepackage{floatrow}
\renewcommand{\floatpagefraction}{.99}
\newfloatcommand{capbtabbox}{table}[][\FBwidth]
\usepackage{eurosym}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{geometry}
\usepackage{bbm}
\usepackage{hyperref}
\usepackage[titletoc]{appendix}
\usepackage{graphicx}
\newcommand{\argmin}{\arg\!\min}
\usepackage{amsmath, amssymb,amsthm,mathtools,dsfont}
\usepackage{algorithmic}
\usepackage{booktabs}
\usepackage{setspace}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage[flushleft]{threeparttable}
\usepackage{rotating}
\usepackage{float}
\usepackage{tabulary}
\usepackage{ragged2e}
\usepackage{rotating}
\usepackage{epstopdf}
\usepackage[labelfont=bf]{caption}
\usepackage{array}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage{subfig}
\usepackage{placeins}
\usepackage{pxfonts}
\usepackage{xcolor}
\usepackage{svg}
\usepackage{multirow}
\usepackage{verbatim}
\usepackage{soul}
\sethlcolor{yellow}
\hypersetup{
	colorlinks,
	linkcolor={red!50!black},
	citecolor={red!50!black},
	urlcolor={blue!80!black}
}
\textheight=23cm \textwidth=16.5cm \oddsidemargin=0cm
\evensidemargin=0cm \topmargin=-1.75cm
\setcounter{MaxMatrixCols}{10}
\makeatletter
\g@addto@macro\@floatboxreset\centering
\makeatother
\setlength\floatsep{2\baselineskip plus 3pt minus 2pt}
\setlength\textfloatsep{2\baselineskip plus 3pt minus 2pt}
\setlength\intextsep{2\baselineskip plus 3pt minus 2 pt}
\newcommand*{\MyIndent}{\hspace*{0.5cm}}%
\usepackage[normalem]{ulem}
\newtheorem{theorem}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{example}{Example}
\newtheorem{lemma}{Lemma}
\newtheorem{remark}{Remark}
\newtheorem{assumption}{Assumptions}
\newtheorem{hypothesis}{Hypothesis}


% Uncomment the next line to use the harvard package with bibtex
%\usepackage[abbr]{harvard}

% This command determines the leading (vertical space between lines) in draft mode
% with 1.5 corresponding to "double" spacing.
%\draftSpacing{1.5}
%\newlength\TableWidth
\usepackage{array}
\newcolumntype{H}{>{\setbox0=\hbox\bgroup}c<{\egroup}@{}}
\newcommand{\Figures}{Figures/}
\newcommand{\Tables}{Tables/}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\title{\textbf{Response to Referee 3 \\ Quantitative Economics MS 2442 \\``Welfare and Spending Effects of \\ Consumption Stimulus Policies''}}
\author{Christopher D. Carroll, Edmund Crawley, William Du, \\ Ivan Frankovic, and H\aa kon Tretvoll}
\date{}

\begin{document}
	\onehalfspacing
	\maketitle
	
	\noindent Thank you for your thoughtful comments and suggestions on our paper ``Welfare and Spending Effects of Consumption Stimulus Policies''. They were all very useful to us in revising the paper. We hope you agree that the paper has improved. In the following, we summarize the main changes we have made based on your, the other referees', and the editor's suggestions. Thereafter, we state each of your comments in italics and provide point-by-point responses to them.
	
	\section{Summary of Main Changes}
	
\input{MainChanges.tex}
	
	\newpage 
	
\section{Comments}
\begin{itemize}
	
	\item \textit{I thank the authors for computing a model without the splurge factor. The results reported in the appendix suggest that such a model performs remarkably well in matching the most important features in the data, with the minor exception of the highest liquid wealth quartile. Therefore, I believe the paper could benefit from a more thorough motivation for including the splurge factor. For instance, is the distribution of $\beta$s in a model without the splurge factor unreasonable? It seems to me that the authors’ stated goal in the abstract (to ``assess the effectiveness of three fiscal stimulus policies'') could be achieved without the somewhat ad-hoc inclusion of the splurge parameter.}
	
	\noindent \textbf{Response.} 
	
	\item \textit{In my previous report, I raised a point regarding the calibration of labor market transition rates across different educational groups. The authors convincingly argue that heterogeneous $E$ to $U$ transitions fit the steady-state data best, as shown by \citet{elsby2010labor}. However, the same reference also indicates that while separation rates vary across educational groups, they are relatively unaffected by business cycles, whereas job-finding rates exhibit significant variation. From the manuscript, it is unclear to me whether the model reflects this characteristic. On page 9, it is stated that ``the employment transition matrix is adjusted so that unemployment remains at the new high level and the expected length of time for an unemployment spell increases,'' but it is unclear to what extent this adjustment	is driven by changes in $E$ to $U$ versus $U$ to $E$. Clarifying this would enhance the reader’s understanding.}	
	
	\noindent \textbf{Response.} 

	\item \textit{I very much appreciate the addition of the full HANK model as a robustness exercise. This	addition effectively demonstrates that the results are not merely an artifact of the partial equilibrium nature of the initial model, while still allowing readers to draw valuable insights from the latter.}

	\noindent \textbf{Response.} Thanks - we agree that the paper is stronger with the inclusion of this exercise. 
\end{itemize}

\bigskip

\noindent Thank you again for your careful advice on our paper. We hope you find our revision satisfactory.

\bibliographystyle{econark}
\bibliography{../../../../HAFiscal.bib,response}
\end{document}
